---
title: "Session 8"
date: "2024-07-18"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme_united: united
    highlight: tango
    code_download: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(dplyr)
library(tidyverse)
library(visdat)
library(knitr)
```

# <span style="color:#61003c;">**Import datasets and present a glimpse, names and header for the four of them. You must have: crunchy_sales_USA,  crunchy_sales_MX, crunchy_wages_USA and crunchy_wages_MX.**</span> 

```{r}
crunchy_sales_MX <- read_excel("CrunchyMX.xlsx", sheet = "Sales")
crunchy_wages_MX <- read_excel("CrunchyMX.xlsx", sheet = "Wages")
```

```{r}
glimpse(crunchy_sales_MX)
```

```{r}
names(crunchy_sales_MX)
```

```{r}
head(crunchy_sales_MX)
```

```{r}
glimpse(crunchy_wages_MX)
```

```{r}
names(crunchy_wages_MX)
```

```{r}
head(crunchy_wages_MX)
```

```{r}
crunchy_sales_USA <- read_excel("CrunchyUSA.xlsx", sheet = "Sales")
crunchy_wages_USA <- read_excel("CrunchyUSA.xlsx", sheet = "Wages")
```

```{r}
glimpse(crunchy_sales_USA)
```

```{r}
names(crunchy_sales_USA)
```

```{r}
head(crunchy_sales_USA)
```

```{r}
glimpse(crunchy_wages_USA)
```

```{r}
names(crunchy_wages_USA)
```

```{r}
head(crunchy_wages_USA)
```

# <span style="color:#61003c;">**Start working with Crunchy_sales_MX. Use vis_miss to identify NA’s.**</span>

```{r}
vis_miss(crunchy_sales_MX)
```

## Look for outliers with a boxplot for numerical variables, replace NA’s with mean if you do not have outliers, replace it with median otherwise. 

```{r}
# Boxplot of CrujiCos, SaladoCos, and BBQCos
boxplot(crunchy_sales_MX$CrujiCos, main = "CrujiCos", col = "lightblue")
boxplot(crunchy_sales_MX$SaladoCos, main = "SaladoCos", col = "lightblue")
boxplot(crunchy_sales_MX$BBQCos, main = "BBQCos", col = "lightblue")
```

```{r}
# Replace NA's with mean
crunchy_sales_MX <- crunchy_sales_MX %>%
  mutate(CrujiCos = ifelse(is.na(CrujiCos), mean(CrujiCos, na.rm = TRUE), CrujiCos),
         SaladoCos = ifelse(is.na(SaladoCos), mean(SaladoCos, na.rm = TRUE), SaladoCos),
         BBQCos = ifelse(is.na(BBQCos), mean(BBQCos, na.rm = TRUE), BBQCos))
```

```{r}
vis_miss(crunchy_sales_MX)
```

## Fill the spaces for month with fill(dataset,.direction=”up or down") Column: Month

```{r}
crunchy_sales_MX <- crunchy_sales_MX %>%
  fill(Month, .direction = "down")
head(crunchy_sales_MX)
```

## Run again the vis_miss at the end to verify you don’t have any NA.

```{r}
vis_miss(crunchy_sales_MX)
```

# <span style="color:#61003c;">**Rename the columns to match the Crunchy_sales_USA names. Show the header of the object.**</span>

```{r}
colnames(crunchy_sales_MX) <- c(
  "Row", "Month", "Region", "CrunchyCo", "SpicyCo", "SaltCo", "BBQCo", "SalsaCo"
)
names(crunchy_sales_MX)
```

# <span style="color:#61003c;">**Add a column ”Country” and write Mexico in every observation. Show the header of the object. **</span>

```{r}
crunchy_sales_MX <- crunchy_sales_MX %>%
  mutate(Country = "Mexico")
head(crunchy_sales_MX)
```

### <span style="color:#61003c;">**5. Repeat steps 2 and 4  for Crunchy_sales_USA (for step 4, write USA).**</span>

```{r}
vis_miss(crunchy_sales_USA)
```

```{r}
boxplot(crunchy_sales_USA$CrunchyCo, main = "CrunchyCo", col = "lightblue")
boxplot(crunchy_sales_USA$SaltCo, main = "SaltCo", col = "lightblue")
boxplot(crunchy_sales_USA$BBQCo, main = "BBQCo", col = "lightblue")
```

```{r}
crunchy_sales_USA <- crunchy_sales_USA %>%
  mutate(CrunchyCo = ifelse(is.na(CrunchyCo), mean(CrunchyCo, na.rm = TRUE), CrunchyCo),
         SaltCo = ifelse(is.na(SaltCo), mean(SaltCo, na.rm = TRUE), SaltCo),
         BBQCo = ifelse(is.na(BBQCo), mean(BBQCo, na.rm = TRUE), BBQCo))
```

```{r}
vis_miss(crunchy_sales_USA)
```

```{r}
crunchy_sales_USA <- crunchy_sales_USA %>%
  fill(Month, .direction = "down")

head(crunchy_sales_USA)
```

```{r}
crunchy_sales_USA <- crunchy_sales_USA %>%
  mutate(Country = "USA")

head(crunchy_sales_USA)
```

```{r}
vis_miss(crunchy_sales_USA)
```

# <span style="color:#61003c;">**Merge with rbind Crunchy_sales_USA and Crunchy_sales_MX and name it crunchy_sales.**</span>

```{r}
crunchy_sales <- rbind(crunchy_sales_USA, crunchy_sales_MX)
```

# <span style="color:#61003c;">**For crunchy_wages_USA and crunchy_wages_MX:**</span>

## In variable “wage”, remove the observation that has text using subset and  change it to numeric.

```{r}
names(crunchy_wages_MX)
# [1] "Number of workers" "Wage (USD)"
```

```{r}
names(crunchy_wages_USA)
# [1] "Number of workers" "Wage (USD)"
```

```{r}
# Más de $62,000 (crunchy_wages_MX) to $62000
crunchy_wages_MX <- crunchy_wages_MX %>%
  subset(!grepl("Más de", `Wage (USD)`)) %>%
  mutate(`Wage (USD)` = as.numeric(gsub("\\$", "", `Wage (USD)`)))

class(crunchy_wages_MX$`Wage (USD)`)
```

```{r}
# More than $41,000 (crunchy_wages_USA) to $41000
crunchy_wages_USA <- crunchy_wages_USA %>%
  subset(!grepl("More than", `Wage (USD)`)) %>%
  mutate(`Wage (USD)` = as.numeric(gsub("\\$", "", `Wage (USD)`)))

class(crunchy_wages_USA$`Wage (USD)`)
```

## Add a column “country” as in step 4.

```{r}
crunchy_wages_MX <- crunchy_wages_MX %>%
  mutate(Country = "Mexico")
head(crunchy_wages_MX)
```

```{r}
crunchy_wages_USA <- crunchy_wages_USA %>%
  mutate(Country = "USA")
head(crunchy_wages_USA)
```

## In crunchy_MX_wages mutate the variable Wage (USD) to change the currency to USD, use 16.84MXN -> 1 USD.

```{r}
crunchy_wages_MX <- crunchy_wages_MX %>%
  mutate(`Wage (USD)` = `Wage (USD)` / 16.84)
head(crunchy_wages_MX)
```

## In crunchy_USA_wages mutate to round up to an integer the variable Number of workers. Hint: decide if you use ceiling or floor function. 

```{r}
# If the decimal part is greater than 0.5, round up, otherwise, round down.
crunchy_wages_USA <- crunchy_wages_USA %>%
  mutate(
    `Number of workers` = ifelse(
      `Number of workers` - floor(`Number of workers`) > 0.5,
      ceiling(`Number of workers`), floor(`Number of workers`)
    )
  )
```

## Merge both datasets into a single one: crunchy_wages.

```{r}
crunchy_wages <- rbind(crunchy_wages_USA, crunchy_wages_MX)
head(crunchy_wages)
```

## For both variables, change the name to avoid white spaces.

```{r}
colnames(crunchy_wages) <- c("NumberOfWorkers", "WageUSD", "Country")
names(crunchy_wages)
```

# <span style="color:#61003c;">**Use kable (library: knitr) to present the 2 final datasets.**</span>

```{r}
kable(crunchy_sales)
```

```{r}
kable(crunchy_wages)
```

```{r, include=FALSE}
# rmarkdown::render(
#   "A01232580_Session8.rmd",
#   output_format = rmarkdown::pdf_document(),
#   output_dir = "./docs/"
# )

# rmarkdown::render(
#   "./src/Session8/A01232580_Session8.rmd",
#   output_format = "html_document",
#   output_dir = "./docs/"
# )
```